---
title: "compare madgetech and arduino reads"
author: "Amelia Ritger"
date: "2022-12-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load libraries
```{r}
library(tidyverse)
library(here)
library(lubridate)
library(scales)

pattern_keep = "1216" #set the date of interest
#start_time <-  "12-16-2022 14:30:00" #to remove 12-15-2022 11:00:00, 12-14-2022 18:00:00, 12-16-2022 14:30:00
#end_time <- "12-16-2022 15:30:00" #to remove 12-15-2022 11:40:00, 12-14-2022 18:30:00, 12-16-2022 15:30:00
```

Load sensor data
```{r}
pattern_remove_deploy = "depl" #exclude deploy files
pattern_remove_calib = "_calib" #exclude calibration files

files_all <- list.files(here("data", "calibrated"), pattern = pattern_keep, recursive = FALSE, full.names = TRUE)
files_tris <- list.files(here("data", "calibrated"), pattern = pattern_remove_deploy, recursive = FALSE, ignore.case = TRUE, full.names = TRUE)
files_calib <- list.files(here("data", "calibrated"), pattern = pattern_remove_calib, recursive = FALSE, full.names = TRUE)
file_names <- setdiff(files_all, files_tris)
file_names <- setdiff(file_names, files_calib)

merged <- read_csv(file_names, col_names = TRUE)

merged_ribbon <- merged %>%
  mutate(tolerance = ifelse(tolerance %in% c("+5", "+1"), "plus", ifelse(tolerance %in% c("-5", "-1"), "minus", "ideal"))) %>%
  #mutate(Logger=ifelse(logger=="arduino", "New design", "Old design")) %>%
  select(1,2,8,12:13) %>%
  group_by(logger) %>%
  filter(!logger=="real-arduino") %>%
  pivot_wider(names_from = tolerance, values_from = c(durafet_temp, omegasph)) %>%
  filter(!date_time==ymd_hms("2022-12-16 16:52:55"),
         !date_time==ymd_hms("2022-12-16 15:31:18")) %>% #remove weird arduino points
  mutate(Logger = ifelse(logger=="arduino", "New design", "Old design"))

ggplot(data=merged_ribbon, aes(x=date_time, group=Logger)) +
  geom_point(aes(y=omegasph_ideal, color=Logger), size=1) +
  labs(x="Date and Time", y="pH") +
  scale_x_datetime(breaks = scales::date_breaks("30 min"), 
                   labels = date_format("%Y-%m-%d %H:%M"))
                   #limits = as.POSIXct(c("2022-12-15 23:00:00","2022-12-16 14:00:00")),
                   #expand = c(0, 0)) +
  scale_color_brewer(palette="Set1", direction=-1) +
  theme_bw() +
  theme(axis.text.x=element_text(angle=20, vjust = 1, hjust=1),
        legend.position = "top")
  
#ggsave(here("figures", "final", "swap-ph.png"), height=10, width=15, units="cm")

ggplot(data=merged_ribbon, aes(x=date_time, group=Logger)) +
  geom_point(aes(y=durafet_temp_ideal, color=Logger)) +
  labs(x="Date and Time", y="Temp (C)") +
  scale_x_datetime(breaks = scales::date_breaks("30 min"), 
                   labels = date_format("%Y-%m-%d %H:%M")) +
                   #limits = as.POSIXct(c("2022-12-15 23:00:00","2022-12-16 14:00:00")),
                   #expand = c(0, 0)) +
  scale_color_brewer(palette="Set1", direction=-1) +
  theme_bw() +
  theme(axis.text.x=element_text(angle=20, vjust = 1, hjust=1),
        legend.position = "top")
  
ggsave(here("figures", "final", "swap-temp.png"), height=10, width=15, units="cm")
```

Calculate difference between swapped sensors 
```{r}
merged_diff <- merged %>%
  mutate(tolerance = ifelse(tolerance %in% c("+5", "+1"), "plus", ifelse(tolerance %in% c("-5", "-1"), "minus", "ideal"))) %>%
  #mutate(Logger=ifelse(logger=="arduino", "New design", "Old design")) %>%
  select(1,2,8,12:13) %>%
  filter(!logger=="real-arduino",
         tolerance=="ideal",
         date_time<ymd_hms("2022-12-16 16:55:15"),
         date_time<ymd_hms("2022-12-16 15:40:08") | date_time>ymd_hms("2022-12-16 16:14:46") ) %>%
  mutate(period = ifelse(date_time>=ymd_hms("2022-12-16 16:00:00"), "2", "1")) %>%
  arrange(date_time) %>%
  group_by(period) %>%
  mutate(diff_time = date_time-lag(date_time, default = first(date_time)),
         diff_temp = ifelse(date_time<ymd_hms("2022-12-16 15:01:39"), durafet_temp-lag(durafet_temp, default=first(durafet_temp)), durafet_temp-lead(durafet_temp, default=first(durafet_temp))),
         diff_ph = ifelse(date_time<ymd_hms("2022-12-16 15:01:39"), omegasph-lag(omegasph, default=first(omegasph)), omegasph-lead(omegasph, default=first(omegasph)))) %>%
  ungroup() %>%
  filter(!date_time==ymd_hms("2022-12-16 15:40:00"),
         !date_time==ymd_hms("2022-12-16 16:55:10"), #remove very last observation
         !date_time==ymd_hms("2022-12-16 14:20:00")) %>% #remove very first observation
  mutate(same = ifelse(logger == lag(logger), "YES", "NO"))

merged_diff_sum <- merged_diff %>%
  group_by(period, logger) %>%
  #mutate(diff_temp = abs(diff_temp),
  #       diff_ph = abs(diff_ph)) %>%
  summarize(mean_temp = mean(durafet_temp),
            mean_diff_t = mean(diff_temp),
            mean_diff_p = mean(diff_ph))

avg1 <- mean(0.1481472,0.1498298) #get these numbers from above
avg2 <- mean(0.1920068,0.1942911) 
avg_temp <- avg2-avg1 #0.0401972 -> #0.0438596

avg3 <- mean(0.2478512,0.2479062)
avg4 <- mean(0.2840733,0.2841385) #get these numbers from above
avg_ph <- avg4-avg3 #0.036221 -> 0.036221
```

Plot differences between swapped sensors
```{r}
ggplot(data=merged_diff, aes(x=date_time, group=logger)) +
  geom_point(aes(y=diff_temp, color=logger), size=1) +
  labs(x="Date and Time", y="Temperature Difference (C)", color = "") +
  scale_x_datetime(breaks = scales::date_breaks("30 min"), 
                   labels = date_format("%Y-%m-%d %H:%M")) +
                   #limits = as.POSIXct(c("2022-12-15 23:00:00","2022-12-16 14:00:00")),
                   #expand = c(0, 0)) 
  scale_y_continuous(limits = c(-0.3, 0.3)) +
  scale_color_brewer(palette="Set1", direction=-1, labels = c("New Design-Old Design", "Old Design-New Design")) +
  #scale_color_discrete(labels=c('label1', 'label2')) +
  theme_bw() +
  theme(axis.text.x=element_text(angle=20, vjust = 1, hjust=1),
        legend.position = "top")

ggsave(here("figures", "final", "swap-temp_diff.png"), height=10, width=15, units="cm")

ggplot(data=merged_diff, aes(x=date_time, group=logger)) +
  geom_point(aes(y=diff_ph, color=logger), size=1) +
  labs(x="Date and Time", y="pH Difference") +
  scale_x_datetime(breaks = scales::date_breaks("30 min"), 
                   labels = date_format("%Y-%m-%d %H:%M")) +
                   #limits = as.POSIXct(c("2022-12-15 23:00:00","2022-12-16 14:00:00")),
                   #expand = c(0, 0)) 
  scale_y_continuous(limits = c(-0.4, 0.4)) +
  scale_color_brewer(palette="Set1", direction=-1, labels = c("New Design-Old Design", "Old Design-New Design")) +
  theme_bw() +
  theme(axis.text.x=element_text(angle=20, vjust = 1, hjust=1),
        legend.position = "top")

ggsave(here("figures", "final", "swap-ph_diff.png"), height=10, width=15, units="cm")
```

Load and merge hobo logger data
```{r}
temp <- read_csv(here("data", "temp", paste(pattern_keep,"_hobo_tris.csv", sep="")), col_names = TRUE) %>%
  select(2:3) %>%
  setNames(c("date_time", "hobo_temp")) %>%
  mutate(date_time=mdy_hms(date_time), #aapply lubridate to date/time column
         logger = "hobo") 
         #date=format(date_time, '%m/%d/%Y'), #create only date column
         #time=format(date_time, '%H:%M:%S')) #create only time column

merged_all <- full_join(merged, temp) %>%
  filter(date_time >= mdy_hms(start_time),
         date_time <= mdy_hms(end_time))
```


Plot up Arduino data comparison
```{r}
merged_ribbon <- merged %>%
  filter(!is.na(logger)) %>%
  select(1,2,8,12,13) %>%
  mutate(tolerance = ifelse(tolerance %in% c("+5", "+1"), "plus", ifelse(tolerance %in% c("-5", "-1"), "minus", "ideal"))) %>%
  pivot_wider(names_from = tolerance, values_from = c(durafet_temp, omegasph))
  #arrange(date_time) %>% filter(date_time <= mdy_hms("12-16-2022 15:00:00"))

ggplot(data=merged_ribbon, aes(x=date_time, group=logger)) +
  geom_point(aes(y=omegasph_ideal, color=logger), size=1)
  #geom_ribbon(aes(ymin=omegasph_minus, ymax=omegasph_plus, fill=logger), alpha=0.2)

ggplot(data=merged_ribbon, aes(x=date_time, group=logger)) +
  geom_point(aes(y=durafet_temp_ideal, color=logger)) +
  geom_ribbon(aes(ymin=durafet_temp_minus, ymax=durafet_temp_plus, fill=logger), alpha=0.2)

ggsave("12-16_not-swapped.png")
```


```{r}
swap <- merged_ribbon %>%
  mutate(time = ifelse(date_time < ymd_hms("2022-12-16 16:00:00"), "pre", "post"),
         sensor = ifelse(time == "pre" & logger == "arduino" | time == "post" & logger == "madgetech" , "1", "2"))

ggplot(data=swap, aes(x=date_time, group=logger)) +
  geom_point(aes(y=omegasph_ideal, color=sensor))
  #geom_ribbon(aes(ymin=omegasph_minus, ymax=omegasph_plus, fill=logger), alpha=0.2)

#0.05 difference in temperature

swap1_p <- swap %>%
  filter(date_time < ymd_hms("2022-12-16 16:00:00")) %>%
  arrange(date_time) %>%
  mutate(diff_time = date_time-lag(date_time, default = first(date_time)),
         diff_temp = omegasph_ideal-lag(omegasph_ideal, default=first(omegasph_ideal))) %>%
    slice(2:n()) %>%
  filter(logger=="arduino") %>%
  summarize(mean_diff = mean(diff_temp),
            mean_temp = mean(omegasph_ideal))

swap2_p <- swap %>%
  filter(date_time > ymd_hms("2022-12-16 16:00:00"),
         date_time < ymd_hms("2022-12-16 16:45:00")) %>%
  arrange(date_time) %>%
  mutate(diff_time = date_time-lag(date_time, default = first(date_time)),
         diff_temp = omegasph_ideal-lag(omegasph_ideal, default=first(omegasph_ideal))) %>%
    slice(2:n()) %>%
  filter(logger=="madgetech") %>%
  summarize(mean_diff = mean(diff_temp),
            mean_temp = mean(omegasph_ideal))
```


```{r}
library(ggpmisc)

merged_comp <- merged_all %>%
  filter(tolerance=="ideal" | is.na(tolerance)) %>%
  mutate(logger = ifelse(is.na(logger), "hobo", logger),
         temp_c = ifelse(logger=="hobo", hobo_temp, durafet_temp))

ggplot(data=merged_comp, aes(x=date_time, y=temp_c, group = logger)) +
  geom_point() +
  stat_poly_line() +
  stat_poly_eq(aes(label = paste(after_stat(eq.label)))) +
  scale_y_continuous(limits=c(18.0, 21.5))

ggsave("temp_corr15.png")
```

Get average difference between hobo and each logger to plot difference across temperature
```{r}
merge_hobo <- merged_all %>%
  filter(tolerance=="ideal" | is.na(tolerance)) %>%
  arrange(date_time) %>%
  mutate(diff = date_time-lag(date_time, default = first(date_time))) %>%
  mutate(temp_down = hobo_temp,
         temp_up = hobo_temp) %>%
  fill(temp_down, .direction = "down") %>%
  fill(temp_up, .direction = "up") %>%
  mutate(hobo_real = ifelse(diff<20,temp_up,temp_down),
         durafet_fill = ifelse(is.na(hobo_temp),durafet_temp, hobo_real)) %>%
  mutate(temp_diff = hobo_real - durafet_temp) %>%
  filter(!is.na(temp_diff)) %>%
  group_by(logger) %>%
  summarize(mean_diff = mean(temp_diff),
            mean_temp = mean(durafet_fill))

hobo_comp_tris_15 <- merge_hobo
```


Get average difference between arduino and madgetech to plot difference across temperature
```{r}
merge_comp <- merged %>%
  filter(tolerance=="ideal" | is.na(tolerance)) %>%
  arrange(date_time) %>%
  mutate(diff_time = date_time-lag(date_time, default = first(date_time)),
         diff_temp = durafet_temp-lag(durafet_temp, default=first(durafet_temp))) %>%
  filter(diff_temp > 0.01 | diff_temp < -0.01) %>%
  mutate(diff_time = date_time-lag(date_time, default = first(date_time)),
         diff_temp = durafet_temp-lag(durafet_temp, default=first(durafet_temp))) %>%
    slice(2:n()) %>%
  filter(logger=="madgetech") %>%
  summarize(mean_diff = mean(diff_temp),
            mean_temp = mean(durafet_temp))
  
  #filter(logger=="madgetech") %>%
  #summarize(mean_diff = mean(diff_temp),
  #          mean_temp = mean(durafet_temp))

  #filter(date_time > ymd_hms("2022-12-15 11:02:40"),
  #       logger=="madgetech") %>%
  #summarize(mean_diff = mean(diff_temp),
  #          mean_temp = mean(durafet_temp))

log_comp_tris_14 <- merge_comp
```
