---
title: "compare madgetech and arduino reads"
author: "Amelia Ritger"
date: "2022-12-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load libraries
```{r}
library(tidyverse)
library(here)
library(lubridate)

pattern_keep = "1215" #set the date of interest
start_time <-  "2022-12-15 13:30:00" #to remove
end_time <- "2022-12-16 13:15:00" #to remove
filter_time <- "2022-12-15 23:00:00" 
```

Load sensor data
```{r}
pattern_remove_tris = "tris" #exclude tris files
pattern_remove_calib = "_calib" #exclude calibration files

files_all <- list.files(here("data", "calibrated"), pattern = pattern_keep, recursive = FALSE, full.names = TRUE)
files_tris <- list.files(here("data", "calibrated"), pattern = pattern_remove_tris, recursive = FALSE, ignore.case = TRUE, full.names = TRUE)
files_calib <- list.files(here("data", "calibrated"), pattern = pattern_remove_calib, recursive = FALSE, full.names = TRUE)
file_names <- setdiff(files_all, files_tris)
file_names <- setdiff(file_names, files_calib)

merged <- read_csv(file_names, col_names = TRUE) #%>% filter(date_time > ymd_hms("2022-12-15 16:00:00" ))
# 
# ggplot(merged, aes(x=date_time, y=durafet_temp)) +
#   geom_point()
```

Load and merge hobo logger data
```{r}
temp <- read_csv(here("data", "temp", paste(pattern_keep,"_hobo_depl.csv", sep="")), col_names = TRUE) %>%
  select(2:3) %>%
  setNames(c("date_time", "hobo_temp")) %>%
  mutate(date_time=mdy_hms(date_time)) #aapply lubridate to date/time column
         #date=format(date_time, '%m/%d/%Y'), #create only date column
         #time=format(date_time, '%H:%M:%S')) #create only time column

merged_all <- full_join(merged, temp)
```

Change sampling range (12/15 - air bubbles for first 8 hours?)
```{r}
merged_sub <- merged %>%
  arrange(date_time) %>%
  filter(date_time >= ymd_hms(filter_time))
```

Plot up Arduino data comparison
```{r}
merged_ribbon <- merged_sub %>%
  filter(!is.na(logger)) %>%
  mutate(tolerance = ifelse(tolerance %in% c("+5", "+1"), "plus", ifelse(tolerance %in% c("-5", "-1"), "minus", "ideal"))) %>%
  select(1:5) %>%
  pivot_wider(names_from = tolerance, values_from = c(durafet_temp, p_h))

#plot up temp data
ggplot(data=merged_ribbon, aes(x=date_time, group=logger)) +
  #geom_point(aes(y=p_h, color=tolerance, shape=logger))
  geom_line(aes(y=durafet_temp_ideal, color=logger)) +
  geom_ribbon(aes(ymin=durafet_temp_minus, ymax=durafet_temp_plus, fill=logger), alpha=0.2) +
  labs(x="Date and Time", y="Temp (C)", label="Logger")

ggsave(here("figures", "deployment-temp_subset.png"), height=20, width=40, units="cm")

#plot up pH data
ggplot(data=merged_ribbon, aes(x=date_time, group=logger)) +
  #geom_point(aes(y=p_h, color=tolerance, shape=logger))
  geom_line(aes(y=p_h_ideal, color=logger)) +
  geom_ribbon(aes(ymin=p_h_minus, ymax=p_h_plus, fill=logger), alpha=0.2) +
  labs(x="Date and Time", y="pH")

ggsave(here("figures", "deployment-ph_subset.png"), height=20, width=40, units="cm")

#look at average difference between madgetech and arduino readings for temp and pH
merged_ribbon_compare <- merged_ribbon %>%
  arrange(date_time) %>%
  mutate(diff_temp = durafet_temp_ideal-lag(durafet_temp_ideal, default=first(durafet_temp_ideal)),
         diff_ph = p_h_ideal-lag(p_h_ideal, default=first(p_h_ideal))) %>%
  slice(2:n()) %>%
  group_by(logger) %>%
  summarize(mean_temp_diff = mean(diff_temp),
            sd_temp_diff = sd(diff_temp),
            mean_ph_diff = mean(diff_ph),
            sd_ph_diff = sd(diff_ph))
```

Plot temperature section to compare between HOBO and sensors
```{r}
library(ggpmisc)

merged_comp <- merged_all %>%
   filter(date_time > ymd_hms(start_time),
          date_time < ymd_hms(end_time)) %>%
  filter(tolerance=="ideal" | is.na(tolerance)) %>%
  mutate(logger = ifelse(is.na(logger), "hobo", logger),
         temp_c = ifelse(logger=="hobo", hobo_temp, durafet_temp))

ggplot(data=merged_comp, aes(x=date_time, y=temp_c, group=logger)) +
  #geom_point(aes(y=p_h, color=tolerance, shape=logger))
  geom_point() +
  stat_poly_line() +
  stat_poly_eq(aes(label = paste(after_stat(eq.label)))) +
  scale_y_continuous(limits=c(7.0, 8.8))

ggsave("temp_correction_deploy.png")
```

Get average difference to plot points at different temperatures
```{r}
merge_hobo <- merged_sub %>%
  filter(date_time > ymd_hms(start_time),
          date_time < ymd_hms(end_time)) %>%
  filter(tolerance=="ideal" | is.na(tolerance)) %>%
  arrange(date_time) %>%
  mutate(diff = date_time-lag(date_time, default = first(date_time))) %>%
  mutate(temp_down = hobo_temp,
         temp_up = hobo_temp) %>%
  fill(temp_down, .direction = "down") %>%
  fill(temp_up, .direction = "up") %>%
  mutate(hobo_real = ifelse(diff<20,temp_up,temp_down),
         durafet_fill = ifelse(is.na(hobo_temp),durafet_temp, hobo_real)) %>%
  filter(!is.na(logger)) %>%
  mutate(temp_diff = hobo_real - durafet_temp) %>%
  group_by(logger) %>%
  summarize(mean_diff = mean(temp_diff),
            mean_temp = mean(durafet_fill))

hobo_comp_deploy <- merge_hobo
```


Get average difference between arduino and madgetech to plot difference across temperature
```{r}
merge_comp <- merged %>%
  filter(tolerance=="ideal" | is.na(tolerance)) %>%
  arrange(date_time) %>%
  mutate(diff_time = date_time-lag(date_time, default = first(date_time)),
         diff_temp = durafet_temp-lag(durafet_temp, default=first(durafet_temp))) %>%
  filter(diff_temp > 0.01 | diff_temp < -0.01) %>%
  mutate(diff_time = date_time-lag(date_time, default = first(date_time)),
         diff_temp = durafet_temp-lag(durafet_temp, default=first(durafet_temp))) %>%
    slice(2:n()) %>%
  filter(logger=="arduino") %>%
  summarize(mean_diff = mean(diff_temp),
            mean_temp = mean(durafet_temp))

log_comp_deploy <- merge_comp
```



